Dashboard section

### 2022-07-31 Report/Draft Dashboard section.Rmd
Updating to the latest version of the dashboard.

**Ca. 11 sider med all koden...** (Data reading hører kanskje hjemme et annet sted i rapporten? Eller kanskje som appendix, hvis vi får dårlig plass?)

- pygments, haddock, and kate are nice code syntax highlighting alternatives to the default (https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/)

@@

Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to the user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded:
```{r dashboard_file_validation, eval=FALSE}
validate(need(tools::file_ext(input$html$datapath) == "html",
                  "Please upload an html file."))
validate(need(tools::file_ext(input$excel$datapath) == "xlsx",
                  "Please upload an Excel file."))
```

@@

@@

The revenue and expenses plots rely on a separate 'maximum' function to create the y scales; this is used so that both are on the same scale, and to derive the upper limit of the scale from the supplied data. The lower limit is derived from the supplied data in each plot in order not to suppress any erroneous totals. (This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function reads the Excel sheet anew in order to make use of the Category column:
```{r dashboard_maximum, eval=FALSE}
# Reactive function to create the maximum value for revenue and expenses plots
maximum <- reactive({
  if (datatest() == TRUE) {
    NA
  } else {
    # reading the Excel sheet anew
    read_excel(input$excel$datapath) |>
      # removing the firm and sub-category columns
      select(-Firm, -`Sub-category`) |>
      # filtering to retain only revenue and expenses entries
      filter(Category == "Revenue" | Category == "Expenses") |>
      # grouping and summarising to calculate maximum value
      group_by(Year, Category) |>
      summarise(
        sum = sum(Value, na.rm = TRUE)
      ) |>
      # Creating list of sums
      pull() |>
      # selecting the highest value
      max()
  }
})
```

@@

The dashboard also allows the user to download a pdf version of the analysis through clicking on a button. A `downloadHandler()` function passes the necessary parameters to the R Markdown document and displays a temporary notification while the file is being created. The function includes temporary file creation arguments that should enable it to function on e.g. shinyapps.io and similar server contexts, but this has not been tested.
```{r}
## Download and parameters -----------------------------------------------------
output$report <- downloadHandler(

>>>>

**Ca. 12 sider med all koden...** (Data reading hører kanskje hjemme et annet sted i rapporten? Eller kanskje som appendix, hvis vi får dårlig plass?)

- pygments, haddock, and kate are nice [code syntax highlighting alternatives](https://www.garrickadenbuie.com/blog/pandoc-syntax-highlighting-examples/) to the default.

@@

Flexdashboard and Shiny do not fail particularly gracefully, so in addition to the information to the user and the information about file types provided to the browser, the Shiny logic executed on the button press includes some basic validations using `validate()` and `tools::file_ext()` which output error messages if the wrong file has been uploaded.

@@

Using these functions, the actual creation of data frames takes place in two stages for each file, partially for data testing and partially to create subsidiary data frames used by some functions:
```{r dashboard_data_frames, eval=FALSE}
# Data frames ================================================================
## Creating data frame from html file-----------------------------------------
html_currency <- reactive({
  # validating file
  validate(need(
    tools::file_ext(input$html$datapath) == "html",
    "Please upload an html file."
  ))

  # Testing and creating data frame
  html_test_create(
    input$html$datapath,
    as.character(input$start), as.character(input$end)
  )
})

htmldata <- reactive({
  read_proff_two(html_currency())
})

## Validating and creating long version of Excel file, if possible------------
excel_test <- reactive({
  # validating file
  validate(need(
    tools::file_ext(input$excel$datapath) == "xlsx",
    "Please upload an Excel file."
  ))

  widen_excel_test(input$excel$datapath)
})
```

@@

The revenue and expenses plots rely on a separate 'maximum' function to create the y scales; this is used so that both are on the same scale, and to derive the upper limit of the scale from the supplied data. The lower limit is derived from the supplied data in each plot in order not to suppress any erroneous totals. (This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function reads the Excel sheet anew in order to make use of the Category column:
```{r dashboard_maximum, eval=FALSE}
## scale maximum--------------------------------------------------------------
# Reactive function to create the maximum value for revenue and expenses plots
# the tests are probably not necessary, but included to be on the safe side
maximum <- reactive({
  if (datatest() == TRUE || setequal(
    names(exceldata()),
    c("year", "firm", "category", "sub_category", "name", "value")
  )) {
    NA
  } else {
    excel_test() |>
      # removing the firm and sub-category columns
      select(-firm, -sub_category) |>
      # filtering to retain only revenue and expenses entries
      filter(category == "Revenue" | category == "Expenses") |>
      # grouping and summarising to calculate maximum value
      group_by(year, category) |>
      summarise(
        sum = sum(value, na.rm = TRUE)
      ) |>
      # Creating list of sums
      pull() |>
      # selecting the highest value
      max()
  }
})
```

@@

The dashboard also allows the user to download a pdf version of the analysis through clicking on a button. A `downloadHandler()` function passes the necessary parameters to the R Markdown document and displays a temporary notification while the file is being created. The function includes temporary file creation arguments that should enable it to function on e.g. shinyapps.io and similar server contexts, but this has not been tested. In order to prevent the user from the downloading a nonsensical html file if bad data have been uploaded or entered, the button creation and report generation are located in the second `observe() |> bindEvent()` structure. The creation of the download button is connected to two convenience logical tests, the primary purpose of which is to connect the button to the `validate()` calls upstream in the reactive chain.
```{r}
# Report Output ================================================================

## Download button -------------------------------------------------------------
output$downloadUI <- renderUI({
  if (available_excel() && available_html()) {
    downloadButton("report", "Download report")
  }
})

## Data exists condition testing------------------------------------------------
# Essentially pretty random tests to connect to the individual
# validate() functions upstream in the reactive chain

# Excel
available_excel <- reactive({
  TRUE
}) |>
  bindEvent(setequal(
    names(exceldata()),
    c("year", "firm", "category", "sub_category", "name", "value")
  ))

# HTML
available_html <- reactive({
  TRUE
}) |> bindEvent("lederlonn" %in% names(htmldata()))
## Download and parameters -----------------------------------------------------
output$report <- downloadHandler(
